THE HYDROGEN EPOCH OF REIONIZATION ARRAY DISH: CHARACTERIZATION OF THE INSTRUMENTAL RESPONSE WITH REFLECTOMETRY MEASUREMENTS AND ITS IMPLICATIONS ON THE MEASUREMENT OF 21CM POWER SPECTRUM.

### 2016-02-16 DishReflectometry/Paper1_v1.tex
Added figure 4,5 from new measurements and rewriting the result section

%\documentclass[12pt,preprint]{aastex}
\documentclass[iop]{emulateapj}
\usepackage{float,amsmath}
%\usepackage[margin=1in]{geometry}
%\usepackage{titlesec} %used to format titles
\usepackage{graphicx} %for handling figures
%\usepackage[none]{hyphenat} %disallows hyphenated words
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\volt}{{v}}
\newcommand{\vis}{{V}}
\newcommand{\sky}{{\rm sky}}
\newcommand{\bmvolt}{{a}}
\newcommand{\beam}{{A}}
\newcommand{\thhat}{{\hat\theta}}
\newcommand{\fngexp}{{e^{\frac{2\pi i\nu\vec{b}\cdot\thhat}{c}}}}
\newcommand{\ifngexp}{{e^{-\frac{2\pi i\nu\vec{b}\cdot\thhat}{c}}}}
\newcommand{\dfngexp}{{e^{2\pi i\nu \Delta \tau}}}
\newcommand{\myemail}{skywalker@galaxy.far.far.away}

%\usepackage{cite}

%\citestyle{aa}
\shorttitle{HERA dish reflectometry }
\shortauthors{Patra et al.}

@@

\subsection{Visibility and delay spectrum computation considering multiple reflections}
Upon first incidence of the sky signal $\volt_{sky}$, $\Gamma_{f}$ fraction of the voltage is reflected off the feed while $(1-\Gamma_{f})$ fraction is coupled to the feed. The reflected voltage then further reflected off the dish and $(1-\Gamma_{f}) $ fraction of it reenters the feed with a roundtrip time delay $\Delta \tau$. Hence, if $I_{sky}$ is reflected n times in between the feed and the dish, the net voltage entering the feed after an
$n^{th}$ reflection off the feed and the dish is written as:
\begin{eqnarray}
\volt_{1} & = &  \bmvolt (1-\Gamma_{f}) \volt_{sky}[1+ \Gamma_{f}\Gamma_{d} \dfngexp \nonumber \\
	&& + (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ \nonumber \\
&&  ....+ (\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}] \nonumber
\end{eqnarray}
and, 
\begin{eqnarray}
\volt_{2} & = &  \bmvolt (1-\Gamma_{f}) \volt_{sky}[1+ \Gamma_{f}\Gamma_{d} \dfngexp \nonumber\\
&&+  (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ \nonumber \\
&& ....+ (\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}] \ast \fngexp \nonumber
\end{eqnarray}
%\begin{eqnarray}
%\volt_{1} & = & a \volt_{sky}(1-\Gamma_{f}) + a \volt_{sky} \Gamma_{f} \Gamma_{d} \dfngexp  (1-\Gamma_{f})^{2}\nonumber\\
   %      & = & a \volt_{sky} (1-\Gamma_{f}) \left [ 1+ (1-\Gamma_{a})\Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]\nonumber
%\end{eqnarray}
%\begin{eqnarray}
%\volt_{2} & = & \left [ a \volt_{sky}(1-\Gamma_{f}) + a \volt_{sky} \Gamma_{f} \Gamma_{d} \dfngexp (1-\Gamma_{f})^{2}\right ] \fngexp \nonumber\\
   %      & = & a \volt_{sky} (1-\Gamma_{f}) \left [ 1+(1-\Gamma_{a}) \Gamma_{f} \Gamma_{d} \dfngexp  \right] \fngexp \nonumber
%\end{eqnarray}
Frequency as well as angular dependence of the quantities are not written explicitly for notational simplicity. We also absorb the common gain term $(1-\Gamma_{f})$ with antenna field pattern $a$ and denote them as $\bmvolt_{f}= \bmvolt(1-\Gamma_{f})$ here after. The cross power generated by $v_{1}$ and $v_{2}$ would have spurious visibility response due to mutual correlation between multiply reflected voltages. Fourier transform of this visibility spectrum along the frequency axis results in the delay spectrum. 
In the delay domain, visibility contributed by any two voltage components from two antennas with no mutual delay will be located at the delay $\tau = 0$ whereas any two voltage components from two antennas components having a mutual delay of $n\Delta \tau$ will be centred at $n\Delta \tau$. These will broaden the width of the delay spectrum which finally interferes with the 21cm power spectrum detection. Combining these terms, the expression of visibility will be,

@@

%\begin{eqnarray}
%\Sigma(\vec b, \nu) & = & \bm_{f} |\left [ 1+ (1-\Gamma_{f}) \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]|^{2}\nonumber\\
 %	    & = &  \bm_{f}  \left[ 1+ |(1-\Gamma_{f})|^{2}|\Gamma_{f}|^{2}|\Gamma_{d}|^{2} + 2 \Re \left ( (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp \right) \right] 
%\label{eq{5}}	    
%\end{eqnarray}

%Instrumental response function acquires additional terms arising due to correlation between the direct and reflected component of the sky signal. Fourier transform of this visibility spectrum along the frequency axis results in the delay spectrum. 
%Fourier transform of the visibility contributed by any two voltage components from two antennas with no mutual delay will be located at the same delay $\tau = 0$ in the Fourier domain whereas any two voltage components from two antennas components having a mutual delay of $\Delta \tau$ will be centred at $\tau+\Delta \tau$. This additional delay extends the upper limit on the maximum geometric delay $\tau_{g}$ to $(\tau_{g}+\Delta \tau)$.
 Thus, the correlation between the reflected signal  from the feed and subsequently from the dish with the direct signal extends the foreground response to delays beyond the maximum geometric delay for a given baseline and contaminates the delay mode where 21~cm delay spectrum could potentially be detected.

 %If the dish vertex is so constructed such that the reflection from the dish vertex is negligible i.e, $\Gamma_{d}<<1\ or = 0$ then the second and the third term of the $\Sigma$ vanish. Also, the second term in the expression of $\Sigma(\vec b, \nu)$ is negligible compared to the first and the third term of the expression and, therefore, can be neglected. \\
%===================================

The design specification of HERA elements requires that any
signal, arriving at the feed at a delay $\Delta \tau$ with respect to the direct incidence, should be
at the level of $-60dB$ at a delay of $60ns$ relative to the first incident
signal at the feed \citep{parsons_deboer_memo}. This specification was
approximated based on the power level of the cosmological signal in relation to
foreground signals, which is estimated to be six orders of magnitude fainter
\citep{santos_et_al2005,ali_et_al2008,deoliveira2008,jelic_et_al2008,bernardi_et_al2009,bernardi_et_al2010,ghosh_et_al2011}. Additionally, the $14m$ HERA baselines set a foreground containing
horizon-limit (the wedge) that, with some buffer, sets a delay specification of
$60ns$
\citep{parsons_et_al2012b,vedantham_et_al2012,nithya_et_al2013,liu_et_al2014a,liu_et_al2014b}. Our goal is to measure the instrument response $\Sigma(\nu)$ and its effect on the power spectrum measurements. 
%%%%%%%%%%%%%%%
\section{Measurements}
\begin{figure*}
\centering
\includegraphics[trim={2cm 10cm 20cm 5cm},clip, totalheight=0.3\textheight]{plots/herafeed.jpg}
\vspace{1.0 em}
\caption{HERA feed consists of a pair of crossed dipole ove xxx mt diameter backplane made of wire mesh. The backplane is surrounded by xxx m wire mesh around the edge resulting in encasing the cross dipole in a cylindrical cage.}
\label{fig:herafeed}
\end{figure*}

We carried out reflectometry measurements at the HERA element prototype in Green
Bank, WV (Figure \ref{fig:heradish}) in order to measure the instrument response of the feed and dish assembly and characterise its performance. An HERA element consists off a $14\,m$ diameter
parabolic reflector and a crossed-dipole pair as a feed. The cross dipole antenna pair is identical to the feed of the PAPER antenna which is suspended at the focal plane of the dish with the support of three vertical poles. HERA elements are closely spaced with centre to centre distance between two dishes are equal to the dish diameter. Therefore, to reduce the coupling between the adjacent dishes, the crossed dipole feed along with the back plane is encased in a cylindrical cage. The feed is raised and
lowered by a three-pulley system mounted on the three poles. The focal height of the dish is $4.5m$
($\sim{14.76}$ft).

%As HERA progresses as an experiment, it is necessary to build optimised dishes that aim to minimise the
%challenges of chromaticity in our quest for the Epoch of Reionization.

As given in the previous section, if the incident power from the sky signal is $I_{sky}$, the feed
reflection coefficient is $\Gamma_{f}$, and the dish reflection
coefficient is $\Gamma_{d}$, then the net power entering a feed after
$n^{th}$ reflection off the feed and the dish is:

@@

%where, $\Delta \tau$ is the roundtrip propagation delay of a frequency $\nu$ due to a reflection over the focal distance $l$.

@@

%%%%%%%%%%%%%%%%%
\subsection{Method}
The instrument response kernel is measured by measuring the return loss of the HERA element at the output port of the feed using a vector network analyser. The vector network analyser transmits a broadband noise of known amplitude and phase via a $50ft$ cable up to the feed. While a part of this signal  returns to the VNA post reflection from the feed due to mismatch between the cable and the antenna impedance resulting in the primary return loss, rest of the signal is radiated out by the antenna and illuminates the dish and radiated into the space. However, the reflected signal from the dish vertex returns to
the feed. This incident signal is now
reflected back and forth in between the feed and the dish much like the sky
signal reflection discussed previously.

@@

%++++++++++++++++++++++++++++++++++++++++++++++++++++++

\subsection{Estimation of the delay spectrum of the instrument}
Frequency domain data, as shown in Figure \ref{fig:window}, is Fourier-transformed to compute the response of the system in the delay domain. The signal transmitted by the VNA is reflected first from the feed. In delay domain,  this appears at the zero delay bin. We subtract this value from the measured data and divide the the data by a factor $(1-\Gamma_{a})$ to derive the delay spectrum for the system operating in receiving mode during sky observation.   %Because the measured data is band-limited between $50$ to $500MHz$, it is analogous to multiply the measured data by a square window function. 
%Then in delay domain, the response would be convolved with a $sinc$ function. This may result in excess power at high
%delays due to the sidelobes of the $sinc$ function. Hence, appropriate windowing of the measured data set is necessary before taking the Fourier transform.
%We have chosen a Hamming window for our analysis, and its effectiveness of this
%window function compared to others is illustrated in Figure
%\ref{fig:window}.

As mentioned in Section \ref{sec:theory}, there is a mismatch in amplitude
between the reflections that we measure (originating from the FieldFox pulse)
and reflections produced by sky signal. The reflections that we measure (at high
delays) must be lowered by a factor to represent weaker reflections that would
occur after most of the sky signal is received by the feed. For our
compensation, we multiply our entire delay spectrum by its DC component, which is the feed reflection coefficient $\Gamma_{f}$, and also divide by ($1-\Gamma_{f}$). In other words, these correction factors equate Equation \ref{eqn:ratio2} with Equation \ref{eqn:ratio1}. 
We note that this correction is only accurate at high delays where our
reflections of interest occur. At low delays, our spectrum amplitude should be
increased to represent the original sky signal, but we do not apply this
correction because it is not relevant to our analysis.

\begin{figure*}
\begin{minipage}[b]{0.5\linewidth}
\centering
\includegraphics[width=\linewidth]{plots1/delayspectrum.png}
\caption{Delay transform of the transfer function of the HERA element. The complex return loss of the element is measured between 50 to 400~MHz when the feed is kept at the focal plane of the dish at a focal height 4.5m. The delay spectrum is produced by taking the Fourier transform of the magnitude of the return loss measurement after limiting the measurements between 100 to 200~MHzusing Blackman-Harris window. The intersection of the horizontal and vertical dashed line shows the instrument performance specification as predicted by the theory of delay transform in order to detect the 21cm power spectrum.}
\label{fig:window}
\end{minipage}
\begin{minipage}[b]{0.5\linewidth}
\centering
\includegraphics[width=\linewidth]{plots1/delayspectrum_feed.png}
\caption{Delay transform of the transfer function of the HERA feed element. The feed is kept on the ground downside up so that the entire radiation from the feed is radiated out. Once again, the measurement is done between 50 to 450 MHz and the delay spectrum is produced between $100-200$~MHz by using a Blackman-Harris window. The black dashed lines illustrate the instrument specification.}
\label{fig:outofthedish}
\end{minipage}
\end{figure*}

\section{Results}
\begin{figure*}
\centering
\includegraphics[totalheight=0.3\textheight]{plots/frequency_amp_phase_fullbw.png}
\caption{Amplitude and phase of the measured return loss. Colored dashed lines
mark three different frequency bands: $140-160MHz$ (typical power spectrum bandwidth),
$100-200MHz$ (PAPER bandwidth), and
$50-250MHz$ (HERA bandwidth).}
\label{fig:freq}
\end{figure*}

Figure \ref{fig:freq} shows the return loss measured between $50$ to
$500MHz$, measured when the feed is suspended at the focal plane of the parabolic dish which is at $5m$ from the dish vertex. 
This measurement is Fourier transformed to obtain the delay spectrum of the instrument in the receiving mode as given in equation \ref{eq3}.

>>>>

\documentclass[twocolumn]{emulateapj}
\usepackage{float,amsmath}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{color}
\usepackage{hyperref}
\citestyle{aa}

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\volt}{{v}}
\newcommand{\vis}{{V}}
\newcommand{\sky}{{\rm sky}}
\newcommand{\bmvolt}{{a}}
\newcommand{\beam}{{A}}
\newcommand{\thhat}{{\hat\theta}}
\newcommand{\fngexp}{{e^{\frac{2\pi i\nu\vec{b}\cdot\thhat}{c}}}}
\newcommand{\ifngexp}{{e^{-\frac{2\pi i\nu\vec{b}\cdot\thhat}{c}}}}
\newcommand{\dfngexp}{{e^{2\pi i\nu \Delta \tau}}}
\newcommand{\myemail}{skywalker@galaxy.far.far.away}

\shorttitle{HERA dish reflectometry }
\shortauthors{Patra et al.}

@@

%\subsection{Visibility and delay spectrum computation considering multiple reflections}
Upon first incidence of the sky signal $\volt_{sky}$, $\Gamma_{f}$ fraction of the voltage is reflected off the feed while $(1-\Gamma_{f})$ fraction is coupled to the feed. The reflected voltage then further reflected off the dish and $(1-\Gamma_{f}) $ fraction of it reenters the feed with a roundtrip time delay $\Delta \tau$. Hence, if $I_{sky}$ is reflected n times in between the feed and the dish, the net voltage entering the feed after an
$n^{th}$ reflection off the feed and the dish is written as:
\begin{eqnarray}
\volt_{1} & = &  \bmvolt (1-\Gamma_{f}) \volt_{sky}[1+ \Gamma_{f}\Gamma_{d} \dfngexp \nonumber \\
	&& + (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ \nonumber \\
&&  ....+ (\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}] \nonumber
\end{eqnarray}
and, 
\begin{eqnarray}
\volt_{2} & = &  \bmvolt (1-\Gamma_{f}) \volt_{sky}[1+ \Gamma_{f}\Gamma_{d} \dfngexp \nonumber\\
&&+  (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ \nonumber \\
&& ....+ (\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}] \ast \fngexp \nonumber
\end{eqnarray}
Frequency as well as angular dependence of the quantities are not written explicitly for notational simplicity. We also absorb the common gain term $(1-\Gamma_{f})$ with antenna field pattern $a$ and denote them as $\bmvolt_{f}= \bmvolt(1-\Gamma_{f})$ here after. The cross power generated by $v_{1}$ and $v_{2}$ would have spurious visibility response due to mutual correlation between multiply reflected voltages. Fourier transform of this visibility spectrum along the frequency axis results in the delay spectrum. 
In the delay domain, visibility contributed by any two voltage components from two antennas with no mutual delay will be located at the delay $\tau = 0$ whereas any two voltage components from two antennas components having a mutual delay of $n\Delta \tau$ will be centred at $n\Delta \tau$. These will broaden the width of the delay spectrum which finally interferes with the 21cm power spectrum detection. Combining these terms, the expression of visibility will be,

@@

 Thus, the correlation between the reflected signal  from the feed and subsequently from the dish with the direct signal extends the foreground response to delays beyond the maximum geometric delay for a given baseline and contaminates the delay mode where 21~cm delay spectrum could potentially be detected. \\
The design specification of HERA elements requires that any
signal, arriving at the feed at a delay $\Delta \tau$ with respect to the direct incidence, should be
at the level of $-60dB$ at a delay of $60ns$ relative to the first incident
signal at the feed \citep{parsons_deboer_memo}. This specification was
approximated based on the power level of the cosmological signal in relation to
foreground signals, which is estimated to be six orders of magnitude fainter
\citep{santos_et_al2005,ali_et_al2008,deoliveira2008,jelic_et_al2008,bernardi_et_al2009,bernardi_et_al2010,ghosh_et_al2011}. Additionally, the $14m$ HERA baselines set a foreground containing
horizon-limit (the wedge) that, with some buffer, sets a delay specification of
$60ns$
\citep{parsons_et_al2012b,vedantham_et_al2012,nithya_et_al2013,liu_et_al2014a,liu_et_al2014b}. Our goal is to measure the instrument response $\Sigma(\nu)$ and its effect on the power spectrum measurements. 
%++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Reflectometry measurements}
\begin{figure*}
\centering
\includegraphics[trim={2cm 10cm 20cm 5cm},clip, totalheight=0.3\textheight]{plots/herafeed.jpg}
\vspace{1.0 em}
\caption{HERA feed consists of a pair of crossed dipole ove xxx mt diameter backplane made of wire mesh. The backplane is surrounded by xxx m wire mesh around the edge resulting in encasing the cross dipole in a cylindrical cage.}
\label{fig:herafeed}
\end{figure*}
We carried out reflectometry measurements at the HERA element prototype in Green
Bank, WV (Figure \ref{fig:heradish}) in order to measure the instrument response of the feed and dish assembly and characterise its performance. An HERA element consists off a $14\,m$ diameter
parabolic reflector and a crossed-dipole pair as a feed. The cross dipole antenna pair is identical to the feed of the PAPER antenna which is suspended at the focal plane of the dish with the support of three vertical poles. HERA elements are closely spaced with centre to centre distance between two dishes are equal to the dish diameter. Therefore, to reduce the coupling between the adjacent dishes, the crossed dipole feed along with the back plane is encased in a cylindrical cage. The feed is raised and
lowered by a three-pulley system mounted on the three poles. The focal height of the dish is $5m$.  
As given in the previous section, if the incident power from the sky signal is $I_{sky}$, the feed
reflection coefficient is $\Gamma_{f}$, and the dish reflection
coefficient is $\Gamma_{d}$, then the net power entering a feed after
$n^{th}$ reflection off the feed and the dish is:

@@

@@

%%%%%%%%%%%%%%%%%
\subsection{Measurement equations}
The instrument response kernel is measured by measuring the return loss of the HERA element at the output port of the feed using a vector network analyser. The vector network analyser transmits a broadband noise of known amplitude and phase via a $50ft$ cable up to the feed. While a part of this signal  returns to the VNA post reflection from the feed due to mismatch between the cable and the antenna impedance resulting in the primary return loss, rest of the signal is radiated out by the antenna and illuminates the dish and radiated into the space. However, the reflected signal from the dish vertex returns to
the feed. This incident signal is now
reflected back and forth in between the feed and the dish much like the sky
signal reflection discussed previously.

@@

%++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Results}
\begin{figure*}[ht]
\begin{minipage}[b]{0.5\linewidth}
\centering
\includegraphics[angle=0, width=\linewidth]{plots/s11_mag_fullband.png}
\end{minipage}
\hspace{0.1cm}
\begin{minipage}[b]{0.5\linewidth}
\centering
\includegraphics[angle=0, width=\linewidth]{plots/s11_ph_fullband.png}
\end{minipage}
\caption{Left: Magnitude of the return loss of power measured in between 50 to 500 MHz using vector network analyser. Right panel shows the corresponding phase.}
\label{fig:freq}       
\end{figure*}

Figure \ref{fig:freq} shows the return loss measured between $50$ to
$500MHz$ when the feed is suspended at the focal plane of the parabolic dish which is at $5m$ from the dish vertex. This measurement is Fourier transformed to obtain the delay spectrum of the instrument in the receiving mode as given in equation \ref{eq3}. Measurement bandwidth is kept much larger than the bandwidth of operation to achieve higher delay resolution.

\subsection{Estimation of the delay spectrum of the instrument}
At first, the frequency domain data is Fourier-transformed to compute the response of the system in the delay domain. The signal transmitted by the VNA is reflected first from the feed. In delay domain, this appears at the zero delay bin. In a second iteration , we subtract this value from the measured data and divide the the data by a factor ${(1-\Gamma_{a})\over \Gamma_{f}}$ to map the delay spectrum of the system from transmitting mode to receiving mode during sky observation.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{plots/delay_spectrum_50_500_blackmanHarris.png}
\caption{Delay spectrum of the HERA element computed by taking the Fourier transform of the measured return loss of power at the feed output. The complex return loss of the element is measured between 50 to 500~MHz when the feed is kept at the focal plane of the dish at a focal height 5m.}
\label{fig:window}
\end{figure}

\subsubsection{Effect of dish and the feed on delay spectrum estimation}
\subsubsection{Effect of feed height on delay spectrum estimation}
\subsubsection{Effect of frequency bandwidth on delay spectrum estimation}
\subsection{Comparison with PAPER antenna}
\section{Comparison with simulations}
\section{Effect of instrument delay spectrum on 21cm power spectrum measurements }
